A sample workflow

### 2025-05-01 analyses/lpiworkflow.qmd
small edits to example workflow + add one sentence to manuscript

---
title: "A sample workflow"
authors: "Victor, Lizzie"
date: "April 2025"
date-format: "MMMM YYYY"
toc: true
number-sections: true
highlight: pygments
execute:
  enabled: true
format:
  html:
    html-math-method: katex
    theme:
      - lux
    embed-resources: true
    code-overflow: wrap
    linkcolor: "#B97C7C"
    page-layout: full
    fig-width: 8
    fig-height: 4
  pdf:
    keep-tex: true
    fig-width: 5.5
    fig-height: 7.5
    code-overflow: wrap
    monofontoptions:
      - Scale=0.5
knitr:
  opts_chunk:
    comment: ''
  opts_knit:
    global.par: TRUE
format-links: false
---

@@

library(rstan)
library(ggplot2)

@@

Now we check additional parameters. We find the model does a fair job at recovering population-level parameters (aka hyperparameters).

@@

We may want to check what the overall and species trends appear to be:

ggplot() +
  geom_pointrange(data = summ_fit["mu_beta1", c('X50.', 'X2.5.', 'X97.5.')],
                  aes(xmin = X2.5., xmax = X97.5., x = X50., y = "Z"), size = 0.25) +
  geom_text(data = summ_fit["mu_beta1", c('X50.', 'X2.5.', 'X97.5.')],
                  aes(x = -0.045, y = "Z", label = "Overall"), size = 3) +
  geom_pointrange(data = species_trends,
                  aes(xmin = X2.5., xmax = X97.5., x = X50., y = genus_species, color = genus_species), 
                  size = 0.25) +
  scale_color_manual(values = hcl.colors(14, palette = "Dark2"),
                     breaks = unique(lpi_subset_rsh$genus_species)[order(unique(lpi_subset_rsh$genus_species))]) +
  lims(x = c(-0.1,0.1)) +
  labs(x = "Trends", color = "Species") +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank())

To see if that might be a good estimate, we need to check how the model predicts the data on finer scales. So we start by examining some species separately, e.g. here the African bush elephant (each color within a country represents a different population):

@@

>>>>

---
title: "A sample workflow"
author:   
  - Victor
  - Jonathan
  - Lizzie
date: "April 2025"
date-format: "MMMM YYYY"
toc: true
number-sections: true
highlight: pygments
execute:
  enabled: true
format:
  html:
    html-math-method: katex
    theme:
      - lux
    embed-resources: true
    code-overflow: wrap
    linkcolor: "#B97C7C"
    page-layout: full
    fig-width: 8
    fig-height: 4
  pdf:
    keep-tex: true
    fig-width: 5.5
    fig-height: 7.5
    code-overflow: wrap
    monofontoptions:
      - Scale=0.5
knitr:
  opts_chunk:
    comment: ''
  opts_knit:
    global.par: TRUE
format-links: false
---

@@

suppressPackageStartupMessages(library(rstan))
library(ggplot2)

@@

Now we check additional parameters. We find the model does a fair job at recovering population-level parameters.

@@

To evaluate our model more precisely, we need to check how the model predicts the data on finer scales. So we start by examining some species separately, e.g. here the African bush elephant (each color within a country represents a different population):

@@

## Inference

Once we are confident with our model (which is not truly the case here), we can look at the parameter values and check what the overall and species trends appear to be:

ggplot() +
  geom_pointrange(data = summ_fit["mu_beta1", c('X50.', 'X2.5.', 'X97.5.')],
                  aes(xmin = X2.5., xmax = X97.5., x = X50., y = "Z"), size = 0.25) +
  geom_text(data = summ_fit["mu_beta1", c('X50.', 'X2.5.', 'X97.5.')],
                  aes(x = -0.045, y = "Z", label = "Overall"), size = 3) +
  geom_pointrange(data = species_trends,
                  aes(xmin = X2.5., xmax = X97.5., x = X50., y = genus_species, color = genus_species), 
                  size = 0.25) +
  scale_color_manual(values = hcl.colors(14, palette = "Dark2"),
                     breaks = unique(lpi_subset_rsh$genus_species)[order(unique(lpi_subset_rsh$genus_species))]) +
  scale_y_discrete(limits=rev) +
  lims(x = c(-0.1,0.1)) +
  labs(x = "Trends", color = "Species") +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank())
